---
title: "Same-Day Release Policies by Patient Population"
subtitle: "Controlled 420-background policy comparison"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show analysis code"
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(scales)
library(knitr)

find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = FALSE)

  repeat {
    target <- file.path(
      current,
      "full_run_summaries",
      "h1_patient_characteristics_pilot",
      "release"
    )

    if (dir.exists(target)) {
      return(current)
    }

    parent <- dirname(current)

    if (identical(parent, current)) {
      stop(
        "Could not find the repository root. Render this QMD from inside ",
        "CUIMC-Appointment-Simulation."
      )
    }

    current <- parent
  }
}

repo_root <- find_repo_root()

results_root <- file.path(
  repo_root,
  "full_run_summaries",
  "h1_patient_characteristics_pilot",
  "release"
)

selected <- read_csv(
  file.path(results_root, "selected_policy_outcomes.csv"),
  show_col_types = FALSE
)

pairwise <- read_csv(
  file.path(results_root, "pairwise_group_deltas.csv"),
  show_col_types = FALSE
)

objective_switch <- read_csv(
  file.path(results_root, "objective_switch_deltas.csv"),
  show_col_types = FALSE
)

policy_labels <- c(
  baseline = "Baseline",
  horizon_only = "Horizon only",
  reservation_only = "Reservation only",
  both_flexible = "Both flexible"
)

policy_order <- unname(policy_labels)

characteristic_labels <- c(
  no_show_sensitivity = "No-show sensitivity",
  balking_sensitivity = "Balking sensitivity",
  cancellation_propensity = "Cancellation propensity",
  joint_delay_sensitivity = "Joint delay sensitivity"
)

characteristic_order <- unname(characteristic_labels)

contrast_order <- c("same", "mild", "strong")
contrast_labels <- c(
  same = "Same",
  mild = "Mild",
  strong = "Strong"
)

reference_labels <- c(
  low = "Low-sensitivity Class 2",
  moderate = "Moderate-sensitivity Class 2"
)

theme_report <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot",
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      axis.title = element_text(face = "bold")
    )
}

fmt_pp <- function(x, digits = 2) {
  paste0(formatC(100 * x, format = "f", digits = digits), " pp")
}

average_selected <- selected |>
  filter(selection_objective == "average_utilization") |>
  mutate(
    policy_label = factor(
      unname(policy_labels[policy]),
      levels = policy_order
    ),
    characteristic_label = factor(
      unname(characteristic_labels[patient_characteristic]),
      levels = characteristic_order
    ),
    contrast_label = factor(
      unname(contrast_labels[contrast_level]),
      levels = unname(contrast_labels[contrast_order])
    ),
    reference_label = factor(
      unname(reference_labels[class2_reference]),
      levels = unname(reference_labels)
    )
  )

baseline_outcomes <- average_selected |>
  filter(policy == "baseline") |>
  select(
    background_id,
    baseline_average_utilization = average_utilization,
    baseline_weighted_served_rate = weighted_utilization,
    baseline_class1_served_rate = class_1_percent_serviced,
    baseline_class2_served_rate = class_2_percent_serviced
  )

average_selected <- average_selected |>
  left_join(baseline_outcomes, by = "background_id") |>
  mutate(
    utilization_gain =
      average_utilization - baseline_average_utilization,
    weighted_served_rate_gain =
      weighted_utilization - baseline_weighted_served_rate,
    class1_served_rate_gain =
      class_1_percent_serviced - baseline_class1_served_rate,
    class2_served_rate_gain =
      class_2_percent_serviced - baseline_class2_served_rate
  ) |>
  group_by(background_id) |>
  mutate(
    best_average_utilization = max(
      average_utilization,
      na.rm = TRUE
    ),
    practically_best =
      average_utilization >= best_average_utilization - 0.005
  ) |>
  ungroup()

average_pairwise <- pairwise |>
  filter(selection_objective == "average_utilization") |>
  mutate(
    characteristic_label = factor(
      unname(characteristic_labels[patient_characteristic]),
      levels = characteristic_order
    ),
    contrast_label = factor(
      unname(contrast_labels[contrast_level]),
      levels = unname(contrast_labels[contrast_order])
    ),
    reference_label = factor(
      unname(reference_labels[class2_reference]),
      levels = unname(reference_labels)
    ),
    first_policy_label = factor(
      unname(policy_labels[first_policy]),
      levels = policy_order
    ),
    second_policy_label = factor(
      unname(policy_labels[second_policy]),
      levels = policy_order
    )
  )

profile_summary <- average_selected |>
  group_by(
    characteristic_label,
    reference_label,
    contrast_label,
    policy_label
  ) |>
  summarise(
    backgrounds = n_distinct(background_id),
    mean_gain = mean(utilization_gain, na.rm = TRUE),
    median_gain = median(utilization_gain, na.rm = TRUE),
    q25_gain = quantile(utilization_gain, 0.25, na.rm = TRUE),
    q75_gain = quantile(utilization_gain, 0.75, na.rm = TRUE),
    practically_best_share = mean(practically_best, na.rm = TRUE),
    .groups = "drop"
  )

plot_profile <- function(characteristic_value) {
  plot_data <- profile_summary |>
    filter(
      as.character(characteristic_label) ==
        characteristic_value
    )

  ggplot(
    plot_data,
    aes(
      x = contrast_label,
      y = median_gain * 100,
      ymin = q25_gain * 100,
      ymax = q75_gain * 100,
      shape = policy_label
    )
  ) +
    geom_hline(yintercept = 0, colour = "grey55") +
    geom_hline(
      yintercept = 0.5,
      linetype = "dashed",
      colour = "grey70"
    ) +
    geom_pointrange(
      position = position_dodge(width = 0.65),
      linewidth = 0.5
    ) +
    facet_wrap(
      ~reference_label,
      scales = "free_y"
    ) +
    labs(
      x = "Class 1 contrast relative to Class 2",
      y = "Median average-utilization gain (pp)",
      shape = "Policy"
    ) +
    theme_report()
}

collapse_group_effect <- function(x) {
  case_when(
    x == "both_groups_gain" ~
      "Both groups gain",

    x == "priority_gains_general_loses" ~
      "Class 1 gains; Class 2 loses",

    x == "priority_loses_general_gains" ~
      "Class 2 gains; Class 1 loses",

    x %in% c(
      "priority_gains_general_tied",
      "priority_tied_general_gains"
    ) ~
      "One group gains; other ties",

    x %in% c(
      "priority_loses_general_tied",
      "priority_tied_general_loses",
      "both_groups_lose"
    ) ~
      "Loss with no supported gain",

    TRUE ~
      "No clear group change"
  )
}

group_effect_levels <- c(
  "Both groups gain",
  "Class 1 gains; Class 2 loses",
  "Class 2 gains; Class 1 loses",
  "One group gains; other ties",
  "Loss with no supported gain",
  "No clear group change"
)
```

## Main findings

- Most policy differences were practical ties.
- Horizon flexibility was most useful for strong no-show and joint-delay profiles.
- Reservations showed their clearest value for balking-sensitive populations.
- Same-day release limited reservation waste when cancellations were higher.
- Demand and Class 1 share strongly changed policy value.

## Simulation setup

The analysis contains:

- **420 backgrounds**
- **21 behavioural profiles**
- **20 clinic contexts per profile**
- **Four policies per background**
- **Five matched seed replications per policy cell**
- **Same-day release of unused reserved slots**

### Policy parameters

| Policy parameter | Baseline | Horizon only | Reservation only | Both flexible |
|---|---|---|---|---|
| Booking horizon \(H\) | Native horizon | Optimized over 2, 6, 10, 14, 18, 22 and 26 days | Native horizon | Optimized over 2, 6, 10, 14, 18, 22 and 26 days |
| Reserved slots per day \(Q\) | 0 | 0 | Optimized from 0 to daily capacity | Optimized from 0 to daily capacity |
| Reservation window \(W\) | None | None | Optimized from 1 day to the native horizon when \(Q>0\) | Optimized from 1 day to the selected horizon when \(Q>0\) |
| Unused reservation release | Not applicable | Not applicable | Released to the pooled capacity on the service day | Released to the pooled capacity on the service day |

The \(Q\)-\(W\) search used:

- Coarse \(Q\) increments of 5
- Fine \(Q\) increments of 1 near the coarse winner
- Coarse \(W\) increments of 2 days
- Fine \(W\) increments of 1 day near the coarse winner
- \(Q=0\) as a valid option

### Problem parameters

| Problem parameter | Values or construction | Treatment |
|---|---|---|
| Demand-to-capacity ratio (`rho`) | 1.0, 1.6, 2.0, 3.0 | Varied; fully crossed with Class 1 share |
| Class 1 share | 0.1, 0.3, 0.5, 0.7, 0.9 | Varied; fully crossed with `rho` |
| Daily capacity | 30 or 50 slots | Controlled by balanced assignment; not fully crossed |
| Native booking horizon | 6, 14 or 22 days | Controlled by balanced assignment; not fully crossed |
| Class 1 arrival rate | `lambda_1 = rho × capacity × Class 1 share` | Derived |
| Class 2 arrival rate | `lambda_2 = rho × capacity × (1 − Class 1 share)` | Derived |
| No-show behaviour | See the behavioural-profile table | Targeted in no-show and joint profiles; otherwise controlled |
| Balking behaviour | See the behavioural-profile table | Targeted in balking and joint profiles; otherwise controlled |
| Cancellation probability | See the behavioural-profile table | Targeted in cancellation profiles; otherwise controlled |
| Non-target behaviour | Same value for both classes | Controlled within each profile |
| Same-day cancellation capability | Enabled | Fixed |
| Reservation release rule | Same-day release | Fixed |
| Threshold capping | Disabled | Fixed |

Capacity and native horizon were assigned deterministically:

- Capacity 30 and 50 each appeared in 10 of the 20 clinic contexts.
- Native horizons 6, 14 and 22 appeared in 7, 7 and 6 contexts.
- Each demand-share pair received one capacity-horizon combination.
- The six possible capacity-horizon combinations were not all tested within every demand-share pair.

### Numerical behavioural profiles

A delay rule is written as:

`(threshold in days, probability at or below threshold, probability above threshold)`

Lower thresholds and higher probabilities indicate greater sensitivity.

| Behavioural family | Class 2 reference | Class 2 values | Class 1: same | Class 1: mild | Class 1: strong |
|---|---|---|---|---|---|
| No-show | Low | (20, 5%, 10%) | (20, 5%, 10%) | (14, 10%, 20%) | (7, 20%, 30%) |
| No-show | Moderate | (14, 10%, 20%) | (14, 10%, 20%) | (7, 20%, 30%) | (4, 30%, 40%) |
| Balking | Low | (22, 5%, 10%) | (22, 5%, 10%) | (16, 10%, 20%) | (9, 20%, 30%) |
| Balking | Moderate | (16, 10%, 20%) | (16, 10%, 20%) | (9, 20%, 30%) | (5, 30%, 40%) |
| Cancellation | Low | 10% | 10% | 20% | 30% |
| Joint: no-show / balking | Low | (20, 5%, 10%) / (22, 5%, 10%) | Same as Class 2 | (14, 10%, 20%) / (16, 10%, 20%) | (7, 20%, 30%) / (9, 20%, 30%) |
| Joint: no-show / balking | Moderate | (14, 10%, 20%) / (16, 10%, 20%) | Same as Class 2 | (7, 20%, 30%) / (9, 20%, 30%) | (4, 30%, 40%) / (5, 30%, 40%) |

### Non-target behaviour held fixed

| Targeted behavioural family | Behaviour held fixed for both classes |
|---|---|
| No-show sensitivity | Balking = (22 days, 10%, 20%); cancellation = 20% |
| Balking sensitivity | No-show = (4 days, 5%, 10%); cancellation = 20% |
| Cancellation propensity | Balking = (16 days, 10%, 20%); no-show = (14 days, 10%, 20%) |
| Joint delay sensitivity | Cancellation = 20%; no-show and balking are both targeted |

“Same” means that the two classes have identical targeted behaviour.

All non-target behavioural parameters are also identical between classes.

Class 1 share is not a behavioural profile:

- It is the proportion of arrivals belonging to Class 1.
- It is varied within every behavioural profile.
- The 21 profiles do not count Class 1 share as an additional profile.

### Matched seeds and paired confidence intervals

Each policy cell was simulated with five seed values.

- The same seed identifiers were used for the policies compared within a background.
- The policies can follow different queue paths after their decisions diverge.
- The comparison still matches Policy A seed 1 with Policy B seed 1, and so on.

For each matched seed:

> **Seed-level policy difference = Policy A outcome − Policy B outcome**

The paired 95% confidence interval was constructed as follows:

- Calculate the five matched seed differences.
- Resample those five differences with replacement.
- Calculate a mean difference for each resample.
- Repeat 2,000 times.
- Use the 2.5th and 97.5th percentiles as the interval endpoints.

It is called **paired** because the resampling uses within-seed policy differences.

Pairing removes variation shared by the two runs and usually gives a cleaner policy comparison.

A change below **0.005**, or 0.5 percentage points, is treated as a practical tie.

## Overall four-policy comparison

### How to read the figure

- Each panel represents one behavioural characteristic.
- The top-left panel represents **no-show sensitivity**.
- It combines six no-show profiles and 20 clinic contexts per profile.
- Each point is the median policy gain across those 120 backgrounds.
- Gain is measured against the baseline from the **same background**.
- The vertical range is the interquartile range.
- The outcome is overall clinic utilization, not Class 1 utilization alone.
- The dashed line marks a gain of **0.5 percentage points**.

```{r}
#| label: fig-overall-four-policy
#| fig-cap: "Median average-utilization gain relative to the same background's baseline. Intervals show the interquartile range, not a confidence interval. All panels use the same y-axis."
#| fig-width: 12
#| fig-height: 7.5

overall_summary <- average_selected |>
  group_by(characteristic_label, policy_label) |>
  summarise(
    median_gain = median(utilization_gain, na.rm = TRUE),
    q25_gain = quantile(utilization_gain, 0.25, na.rm = TRUE),
    q75_gain = quantile(utilization_gain, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  overall_summary,
  aes(
    x = policy_label,
    y = median_gain * 100,
    ymin = q25_gain * 100,
    ymax = q75_gain * 100
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    colour = "grey70"
  ) +
  geom_pointrange(linewidth = 0.6) +
  facet_wrap(
    ~characteristic_label,
    ncol = 2
  ) +
  labs(
    x = NULL,
    y = "Average-utilization gain (percentage points)"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```


### Baseline utilization

```{r}
#| label: tbl-baseline-utilization
#| tbl-cap: "Baseline average utilization by behavioural family."

baseline_utilization_table <- average_selected |>
  filter(policy == "baseline") |>
  group_by(characteristic_label) |>
  summarise(
    Backgrounds = n_distinct(background_id),
    median_utilization = median(average_utilization, na.rm = TRUE),
    q25_utilization = quantile(average_utilization, 0.25, na.rm = TRUE),
    q75_utilization = quantile(average_utilization, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) |>
  transmute(
    `Behavioural family` = characteristic_label,
    Backgrounds,
    `Median baseline utilization` =
      percent(median_utilization, accuracy = 0.1),
    `Interquartile range` =
      paste0(
        percent(q25_utilization, accuracy = 0.1),
        "–",
        percent(q75_utilization, accuracy = 0.1)
      )
  )

kable(
  baseline_utilization_table,
  align = c("l", "r", "r", "r")
)
```

### Why are the median gains small?

- Baseline utilization was already high in many backgrounds.
- Many low-demand and symmetric-profile backgrounds had almost no recoverable utilization loss.
- The median combines contexts where the policy helped with contexts where it did not.
- Class 1 sometimes represented only 10% or 30% of arrivals.
- A large Class 1 improvement can therefore produce a small change in overall utilization.
- Reservation policies often redistributed service between classes rather than creating many additional completed appointments.

Joint sensitivity does not imply an additive gain:

- Balking and no-show are competing outcomes.
- A patient who rejects an offered delay cannot later no-show for that appointment.
- Shortening the horizon can address both mechanisms, but it cannot recover the same patient twice.
- The median also includes same, mild and strong contrasts under both Class 2 references.
- Some individual joint-sensitivity backgrounds had gains above one percentage point.
- The median across the entire family remained below one percentage point.

### How often was the gain clear?

The next figure compares each active policy directly with baseline.

```{r}
#| label: fig-baseline-evidence
#| fig-cap: "Evidence classification for average-utilization changes relative to baseline. Each bar sums to 100% of backgrounds within a behavioural characteristic and policy."
#| fig-width: 13
#| fig-height: 8

evidence_levels <- c(
  "Supported increase",
  "Meaningful but uncertain increase",
  "Practical tie"
)

baseline_evidence <- average_pairwise |>
  filter(second_policy == "baseline") |>
  mutate(
    evidence = recode(
      average_utilization_change_status,
      supported_increase = "Supported increase",
      uncertain_increase = "Meaningful but uncertain increase",
      practical_tie = "Practical tie",
      supported_decrease = "Supported decrease",
      uncertain_decrease = "Meaningful but uncertain decrease"
    ),
    evidence = factor(
      evidence,
      levels = c(
        evidence_levels,
        "Meaningful but uncertain decrease",
        "Supported decrease"
      )
    )
  ) |>
  count(
    characteristic_label,
    first_policy_label,
    evidence,
    name = "Backgrounds"
  ) |>
  group_by(
    characteristic_label,
    first_policy_label
  ) |>
  mutate(
    share = Backgrounds / sum(Backgrounds)
  ) |>
  ungroup()

ggplot(
  baseline_evidence,
  aes(
    x = first_policy_label,
    y = share,
    fill = evidence
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic_label,
    ncol = 2
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of backgrounds",
    fill = "Result relative to baseline"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

The categories answer different questions:

- **Supported increase**
  - Mean gain is at least 0.005.
  - The 95% interval defined above is above zero.
  - The gain is both practically meaningful and statistically supported.

- **Meaningful but uncertain increase**
  - Mean gain is at least 0.005.
  - The confidence interval still includes zero.
  - More replication may be needed.

- **Practical tie**
  - The absolute mean difference is below 0.005.
  - The difference is too small to treat as operationally important.
  - Statistical significance is not used to promote a very small effect.

This figure conveys **how consistently** a policy helped across backgrounds.

It does not convey the size of the gain. The preceding figure shows magnitude.

### How to read the profile figures

The next four figures use the same summary structure:

- Each point is the median across **20 clinic contexts**.
- A point is specific to one Class 2 reference, one contrast level and one policy.
- Those 20 contexts vary demand, Class 1 share, capacity and native horizon.
- The vertical line is the **interquartile range** across the 20 contexts.
- It is not a confidence interval.

## No-show sensitivity

```{r}
#| label: fig-no-show-profiles
#| fig-cap: "Median policy gain across 20 clinic contexts for each no-show reference and contrast. Lines show interquartile ranges."
#| fig-width: 11
#| fig-height: 5.5

plot_profile("No-show sensitivity")
```

- Under the **low Class 2 reference**, policy gains remained small.
- Under the **moderate reference**, strong Class 1 sensitivity increased the value of shorter horizons.
- In the moderate/strong profile:
  - Horizon only averaged about **0.50 pp** above baseline.
  - Both flexible averaged about **0.55 pp**.
  - Reservation only averaged about **0.05 pp**.
- Both flexible often selected \(Q=0\) in the strong profiles.
- Main implication: **shorter horizons addressed the no-show mechanism more directly than reservations.**

::: {.callout-tip}
**Why can reservations affect no-show outcomes?**

- A reservation does not change either class's no-show rule.
- It can give Class 1 access to a nearer appointment.
- If the new delay moves from above the no-show threshold to at or below it, the assigned no-show probability falls from the high value to the low value.
- If the delay remains on the same side of the threshold, the assigned probability does not change.
- The realized no-show rate can still move irregularly because the policy changes which patients obtain appointments.
:::

::: {.callout-note}
**The “same” profiles are negative controls.**

- Class 1 and Class 2 have identical cancellation, balking and no-show parameters.
- Reservation-only median gains were about **0.08 pp** under the low reference.
- They were about **0.15 pp** under the moderate reference.
- All 40 reservation-only comparisons with baseline were practical ties.
- These small positive points should not be treated as meaningful evidence of a reservation benefit.
- Finite-run variation and selecting the best \(Q\)-\(W\) cell from many candidates can move the selected point estimate slightly above zero.
:::

## Balking sensitivity

```{r}
#| label: fig-balking-profiles
#| fig-cap: "Median policy gain across 20 clinic contexts for each balking reference and contrast. Lines show interquartile ranges."
#| fig-width: 11
#| fig-height: 5.5

plot_profile("Balking sensitivity")
```

- Balking profiles produced the largest reservation-related gains.
- Both flexible averaged approximately **0.6–0.8 pp** above baseline across the six profiles.
- Reservation only usually performed similarly to horizon only in utilization.
- The two policies worked through different mechanisms:
  - Horizon only changed the set of delays offered.
  - Reservations protected near-term capacity for Class 1.
- Stronger sensitivity did **not** create a perfectly monotonic increase.
- Demand and Class 1 share still mattered.

## Cancellation propensity

```{r}
#| label: fig-cancellation-profiles
#| fig-cap: "Median policy gain across 20 clinic contexts as Class 1 cancellation probability increases. Lines show interquartile ranges."
#| fig-width: 10
#| fig-height: 5

plot_profile("Cancellation propensity")
```

- Same-day release prevented a broad utilization penalty from unused reservations.
- Reservation only averaged approximately:
  - **0.31 pp** when cancellation rates were equal.
  - **0.25 pp** under the mild contrast.
  - **0.22 pp** under the strong contrast.
- Both flexible followed the same pattern.
- Higher cancellation reduced reservation value modestly.
- It did not make reservation policies generally worse than baseline.

## Joint delay sensitivity

```{r}
#| label: fig-joint-profiles
#| fig-cap: "Median policy gain across 20 clinic contexts for each joint-sensitivity reference and contrast. Lines show interquartile ranges."
#| fig-width: 11
#| fig-height: 5.5

plot_profile("Joint delay sensitivity")
```

### Why are most effects close to zero?

```{r}
#| label: tbl-joint-delay-exposure
#| tbl-cap: "Baseline delay exposure and policy gains in the joint-sensitivity profiles."

joint_baseline_exposure <- average_selected |>
  filter(
    as.character(characteristic_label) == "Joint delay sensitivity",
    policy == "baseline"
  ) |>
  group_by(
    reference_label,
    contrast_label,
    noshow_threshold_1,
    balk_threshold_1
  ) |>
  summarise(
    median_baseline_delay =
      median(class_1_mean_offered_booking_delay, na.rm = TRUE),
    .groups = "drop"
  )

joint_horizon_summary <- average_selected |>
  filter(
    as.character(characteristic_label) == "Joint delay sensitivity",
    policy == "horizon_only"
  ) |>
  group_by(reference_label, contrast_label) |>
  summarise(
    median_horizon_gain =
      100 * median(utilization_gain, na.rm = TRUE),
    .groups = "drop"
  )

joint_both_summary <- average_selected |>
  filter(
    as.character(characteristic_label) == "Joint delay sensitivity",
    policy == "both_flexible"
  ) |>
  group_by(reference_label, contrast_label) |>
  summarise(
    median_both_gain =
      100 * median(utilization_gain, na.rm = TRUE),
    meaningful_contexts =
      sum(utilization_gain >= 0.005, na.rm = TRUE),
    .groups = "drop"
  )

joint_exposure_table <- joint_baseline_exposure |>
  left_join(
    joint_horizon_summary,
    by = c("reference_label", "contrast_label")
  ) |>
  left_join(
    joint_both_summary,
    by = c("reference_label", "contrast_label")
  ) |>
  transmute(
    `Class 2 reference` = reference_label,
    `Class 1 contrast` = contrast_label,
    `Median baseline Class 1 offered delay` =
      paste0(
        formatC(
          median_baseline_delay,
          format = "f",
          digits = 2
        ),
        " days"
      ),
    `Class 1 no-show / balk threshold` =
      paste0(
        noshow_threshold_1,
        " / ",
        balk_threshold_1,
        " days"
      ),
    `Median horizon-only gain` =
      paste0(
        formatC(
          median_horizon_gain,
          format = "f",
          digits = 2
        ),
        " pp"
      ),
    `Median both-flexible gain` =
      paste0(
        formatC(
          median_both_gain,
          format = "f",
          digits = 2
        ),
        " pp"
      ),
    `Both-flexible contexts ≥0.5 pp` =
      paste0(meaningful_contexts, " of 20")
  )

kable(
  joint_exposure_table,
  align = c("l", "l", "r", "r", "r", "r", "r")
)
```

- Most baseline offers were already shorter than the Class 1 thresholds.
- The policies therefore changed only the tail of long-delay offers.
- In the strongest moderate profile:
  - No-show probability is still 30% at short delays.
  - Balking probability is still 30% at short delays.
  - Moving an offer below the threshold changes 40% to 30%; it does not reduce the probability to zero.
- Balking and no-show are competing outcomes.
  - A patient who balks never reaches the no-show stage.
  - The two potential improvements cannot simply be added.
- Overall utilization counts completed slots across both classes.
  - Avoiding a Class 1 balk may change who receives a slot without increasing the number of completed slots.
  - A utilization gain appears only when the policy prevents a slot from ultimately going unused.
- Large gains were concentrated in a minority of congested contexts.
- When more than half of the 20 context-level gains are very small, the median is also very small.

The clearest response occurred in the moderate/strong profile.

- Horizon flexibility shortened the delay distribution.
- Reservations added value only in selected contexts.
- Both flexible combined those mechanisms.
- The result was context-specific rather than universal.

## Priority-group share

```{r}
#| label: fig-priority-share
#| fig-cap: "Four-policy comparison by Class 1 share. Points show median gains; intervals show the interquartile range across the 21 patient profiles and four demand levels."
#| fig-width: 11
#| fig-height: 5.5

share_summary <- average_selected |>
  group_by(class1_share, policy_label) |>
  summarise(
    median_gain = median(utilization_gain, na.rm = TRUE),
    q25_gain = quantile(utilization_gain, 0.25, na.rm = TRUE),
    q75_gain = quantile(utilization_gain, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  share_summary,
  aes(
    x = factor(class1_share),
    y = median_gain * 100,
    ymin = q25_gain * 100,
    ymax = q75_gain * 100,
    shape = policy_label
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    colour = "grey70"
  ) +
  geom_pointrange(
    position = position_dodge(width = 0.65),
    linewidth = 0.5
  ) +
  labs(
    x = "Class 1 share",
    y = "Median average-utilization gain (pp)",
    shape = "Policy"
  ) +
  theme_report()
```

- Reservation only was most useful at Class 1 shares of **0.1–0.5**.
- Its mean gain fell to about **0.11 pp** at a Class 1 share of 0.9.
- Horizon-only value did not fall in the same way.
- A large Class 1 population already occupies much of the pooled capacity.
- Additional priority protection then provides less incremental value.

## Clinic context

The next figure shows where each active policy improved utilization.

- Each cell is a grouped median.
- It aggregates over patient profiles, capacities and native horizons.
- It should be read as a pattern map, not as a controlled one-variable trend.

```{r}
#| label: fig-demand-share-heatmap
#| fig-cap: "Median average-utilization gain by demand-to-capacity ratio, Class 1 share and active policy."
#| fig-width: 12
#| fig-height: 7

context_summary <- average_selected |>
  filter(policy != "baseline") |>
  group_by(
    policy_label,
    rho,
    class1_share
  ) |>
  summarise(
    median_gain_pp =
      100 * median(utilization_gain, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    rho = factor(rho),
    class1_share = factor(class1_share)
  )

fill_limit <- quantile(
  abs(context_summary$median_gain_pp),
  0.95,
  na.rm = TRUE
)

ggplot(
  context_summary,
  aes(
    x = class1_share,
    y = rho,
    fill = median_gain_pp
  )
) +
  geom_tile(
    colour = "white",
    linewidth = 0.3
  ) +
  geom_text(
    aes(
      label = formatC(
        median_gain_pp,
        format = "f",
        digits = 1
      )
    ),
    size = 3
  ) +
  facet_wrap(
    ~policy_label,
    ncol = 3
  ) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "white",
    high = "#2166AC",
    midpoint = 0,
    limits = c(-fill_limit, fill_limit),
    oob = squish,
    name = "Median gain\n(pp)"
  ) +
  labs(
    x = "Class 1 share",
    y = "Demand-to-capacity ratio"
  ) +
  theme_report()
```

- Policy value was generally low under light demand.
- Differences became more visible under congestion.
- Reservation value depended strongly on the size of Class 1.
- Both flexible was most useful where horizon and reservation mechanisms both remained active.

## Effects on Class 1 and Class 2

Patient-group effects remain comparisons with baseline.

```{r}
#| label: fig-group-effects
#| fig-cap: "Confidence-supported Class 1 and Class 2 served-rate patterns for each active policy compared with baseline."
#| fig-width: 13
#| fig-height: 8

group_effect_summary <- average_pairwise |>
  filter(second_policy == "baseline") |>
  mutate(
    group_effect = factor(
      collapse_group_effect(
        patient_group_supported_tradeoff
      ),
      levels = group_effect_levels
    )
  ) |>
  count(
    characteristic_label,
    first_policy_label,
    group_effect,
    name = "n"
  ) |>
  group_by(
    characteristic_label,
    first_policy_label
  ) |>
  mutate(share = n / sum(n)) |>
  ungroup()

ggplot(
  group_effect_summary,
  aes(
    x = first_policy_label,
    y = share,
    fill = group_effect
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic_label,
    ncol = 2
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of backgrounds",
    fill = "Supported group pattern"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

- Horizon only usually left both groups tied.
- Reservation policies often increased Class 1 served rate while reducing Class 2 served rate.
- This pattern was strongest in balking and symmetric-reference profiles.
- An average-utilization gain should not be described as a gain shared by both groups.

“Loss with no supported gain” combines:

- Class 1 loses while Class 2 ties.
- Class 2 loses while Class 1 ties.
- Both groups lose.

It excludes explicit gain-loss trade-offs.

## Selected policy parameters

### Booking horizon

The figure includes the two policies that optimize \(H\).

```{r}
#| label: fig-horizon-selections
#| fig-cap: "Distribution of selected booking horizons for horizon only and both flexible."
#| fig-width: 12
#| fig-height: 7

horizon_distribution <- average_selected |>
  filter(
    policy %in% c("horizon_only", "both_flexible")
  ) |>
  count(
    characteristic_label,
    policy_label,
    selected_horizon_days,
    name = "n"
  ) |>
  group_by(
    characteristic_label,
    policy_label
  ) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  mutate(
    selected_horizon_days = factor(
      selected_horizon_days,
      levels = c(2, 6, 10, 14, 18, 22, 26)
    )
  )

ggplot(
  horizon_distribution,
  aes(
    x = policy_label,
    y = share,
    fill = selected_horizon_days
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic_label,
    ncol = 2
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of backgrounds",
    fill = "Selected horizon\n(days)"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 20,
      hjust = 1
    )
  )
```

- Short horizons dominated no-show and joint-delay profiles.
- Balking profiles also commonly selected two days.
- Cancellation profiles retained more medium-length horizons.
- The result reflects the discrete candidate grid.

### Reservation quantity and window

The figure reports the optimized \(Q\) and \(W\) values defined in the policy-parameter table.

```{r}
#| label: fig-reservation-parameters
#| fig-cap: "Selected reservation quantity and window for reservation only and both flexible. Points show medians; intervals show interquartile ranges."
#| fig-width: 13
#| fig-height: 8

reservation_long <- bind_rows(
  average_selected |>
    filter(
      policy %in% c(
        "reservation_only",
        "both_flexible"
      )
    ) |>
    transmute(
      characteristic_label,
      policy_label,
      metric = "Reserved quantity (% of capacity)",
      value = 100 * selected_Q / slots_per_day
    ),

  average_selected |>
    filter(
      policy %in% c(
        "reservation_only",
        "both_flexible"
      ),
      selected_Q > 0,
      selected_window >= 1
    ) |>
    transmute(
      characteristic_label,
      policy_label,
      metric = "Positive-Q reservation window (days)",
      value = selected_window
    )
)

reservation_summary <- reservation_long |>
  group_by(
    characteristic_label,
    policy_label,
    metric
  ) |>
  summarise(
    median = median(value, na.rm = TRUE),
    q25 = quantile(value, 0.25, na.rm = TRUE),
    q75 = quantile(value, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  reservation_summary,
  aes(
    x = policy_label,
    y = median,
    ymin = q25,
    ymax = q75
  )
) +
  geom_pointrange(linewidth = 0.6) +
  facet_grid(
    metric ~ characteristic_label,
    scales = "free_y"
  ) +
  labs(
    x = NULL,
    y = "Selected parameter"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

- Balking profiles selected the largest reservations.
- Strong no-show and joint profiles often selected small \(Q\) or no reservation.
- Reservation windows were usually short when the selected horizon was short.
- \(Q\), window and horizon should be interpreted jointly.

## Priority-weighted served rate

The secondary objective is:

> **Priority-weighted served rate = (2 × Class 1 served rate + Class 2 served rate) ÷ 3**

Class 1 receives twice the policy weight of Class 2.

The code calls this metric `weighted_utilization`.

### What changed when the optimization objective changed?

```{r}
#| label: tbl-objective-switch
#| tbl-cap: "Effect of switching from average-utilization-optimal to priority-weighted-optimal policy cells."

switch_summary <- objective_switch |>
  mutate(
    policy_label = factor(
      unname(policy_labels[policy]),
      levels = policy_order
    )
  ) |>
  group_by(policy_label) |>
  summarise(
    Backgrounds = n_distinct(background_id),
    `Same policy cell` =
      percent(mean(same_policy_cell, na.rm = TRUE), accuracy = 0.1),
    `Median average-utilization change` =
      fmt_pp(
        median(delta_average_utilization, na.rm = TRUE)
      ),
    `Median weighted-served-rate change` =
      fmt_pp(
        median(delta_weighted_utilization, na.rm = TRUE)
      ),
    `Median Class 1 served-rate change` =
      fmt_pp(
        median(delta_class_1_percent_serviced, na.rm = TRUE)
      ),
    `Median Class 2 served-rate change` =
      fmt_pp(
        median(delta_class_2_percent_serviced, na.rm = TRUE)
      ),
    .groups = "drop"
  ) |>
  rename(Policy = policy_label)

kable(
  switch_summary,
  align = c("l", "r", "r", "r", "r", "r", "r")
)
```

- Many backgrounds selected the same policy cell under both objectives.
- Where selections changed:
  - Class 1 served rate generally increased.
  - Class 2 served rate often decreased.
  - Average utilization sometimes fell.
- The weighted objective changes **who is served**, not only how much capacity is used.

```{r}
#| label: fig-objective-switch
#| fig-cap: "Average-utilization and priority-weighted-served-rate changes when switching to the priority-weighted-optimal policy cell."
#| fig-width: 10
#| fig-height: 5.5

switch_long <- objective_switch |>
  mutate(
    policy_label = factor(
      unname(policy_labels[policy]),
      levels = policy_order
    )
  ) |>
  select(
    policy_label,
    delta_average_utilization,
    delta_weighted_utilization
  ) |>
  pivot_longer(
    cols = starts_with("delta_"),
    names_to = "metric",
    values_to = "change"
  ) |>
  mutate(
    metric = recode(
      metric,
      delta_average_utilization =
        "Average utilization",
      delta_weighted_utilization =
        "Priority-weighted served rate"
    )
  )

ggplot(
  switch_long,
  aes(
    x = policy_label,
    y = change * 100
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_boxplot(
    outlier.alpha = 0.2,
    width = 0.65
  ) +
  facet_wrap(
    ~metric,
    scales = "free_y"
  ) +
  labs(
    x = NULL,
    y = "Change after objective switch (pp)"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

## Conclusions

| Patient characteristic | Main result | Most relevant comparison |
|---|---|---|
| More no-show sensitive | Horizon flexibility was the main useful mechanism | Horizon only vs both flexible |
| More balking sensitive | Reservations showed their strongest value | Reservation only vs horizon only |
| More cancellation prone | Day-of release limited reservation waste | Reservation only vs baseline |
| Jointly delay sensitive | Strong moderate-reference profiles favoured short horizons | Horizon only vs both flexible |
| Priority-group share | Reservation value declined when Class 1 was already very large | Reservation only across shares |

### What the results support

- Retain all four policy regimes.
- Retain same-day release only.
- Retain the same, mild and strong contrasts.
- Fully cross capacity and native horizon in the confirmatory run.
- Preserve the full Class 1 share range.
- Use this report structure for the final document.
- Treat practical ties as findings, not failures to select a winner.

### What requires confirmation

- Exact effect sizes.
- Capacity-specific conclusions.
- Native-horizon-specific conclusions.
- Fine interactions among profile, demand and share.
- Confirmatory statements about patient-group trade-offs.
