---
title: "Same-Day Release Policies by Patient Population"
subtitle: "Exploratory four-policy comparison using the existing 840 simulation backgrounds"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show analysis code"
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(scales)
library(knitr)
library(purrr)

find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = FALSE)

  repeat {
    marker <- file.path(
      current,
      "full_run_summaries",
      "h1_policy_outcomes",
      "release"
    )

    if (dir.exists(marker)) {
      return(current)
    }

    parent <- dirname(current)

    if (identical(parent, current)) {
      stop(
        "Could not find the repository root. Render this file from inside ",
        "CUIMC-Appointment-Simulation."
      )
    }

    current <- parent
  }
}

resolve_column <- function(data, candidates, pattern = NULL, description) {
  direct <- intersect(candidates, names(data))

  if (length(direct) > 0) {
    return(direct[[1]])
  }

  if (!is.null(pattern)) {
    pattern_hits <- grep(
      pattern,
      names(data),
      value = TRUE,
      ignore.case = TRUE
    )

    if (length(pattern_hits) == 1) {
      return(pattern_hits[[1]])
    }
  }

  stop(
    "Could not identify the ", description, " column. Available columns: ",
    paste(names(data), collapse = ", ")
  )
}

repo_root <- find_repo_root()

release_root <- file.path(
  repo_root,
  "full_run_summaries",
  "h1_policy_outcomes",
  "release"
)

policy_outcomes <- read_csv(
  file.path(release_root, "selected_policy_outcomes.csv"),
  show_col_types = FALSE
)

seed_outcomes <- read_csv(
  file.path(release_root, "selected_policy_seed_outcomes.csv"),
  show_col_types = FALSE
)

pairwise <- read_csv(
  file.path(release_root, "pairwise_group_deltas.csv"),
  show_col_types = FALSE
)

average_metric_col <- resolve_column(
  seed_outcomes,
  candidates = c(
    "average_utilization",
    "average_utilization_value",
    "mean_average_utilization",
    "average_utilization_mean",
    "avg_utilization"
  ),
  pattern = "^average.*utili[sz]ation$|^avg.*utili[sz]ation$",
  description = "seed-level average utilization"
)

class1_served_col <- resolve_column(
  seed_outcomes,
  candidates = c(
    "served_rate_1",
    "class1_served_rate",
    "class_1_served_rate",
    "served_rate_class1",
    "service_rate_1"
  ),
  pattern = "(served|service).*rate.*(_1|class.?1)$|class.?1.*(served|service).*rate",
  description = "Class 1 served rate"
)

class2_served_col <- resolve_column(
  seed_outcomes,
  candidates = c(
    "served_rate_2",
    "class2_served_rate",
    "class_2_served_rate",
    "served_rate_class2",
    "service_rate_2"
  ),
  pattern = "(served|service).*rate.*(_2|class.?2)$|class.?2.*(served|service).*rate",
  description = "Class 2 served rate"
)

seed_col <- resolve_column(
  seed_outcomes,
  candidates = c("seed", "seed_id", "simulation_seed"),
  pattern = "^seed(_id)?$|simulation.*seed",
  description = "seed identifier"
)

required_background_columns <- c(
  "background_id",
  "rho",
  "class1_share",
  "slots_per_day",
  "cancel_1",
  "cancel_2",
  "balk_threshold_1",
  "balk_threshold_2",
  "balk_low_1",
  "balk_low_2",
  "balk_high_1",
  "balk_high_2",
  "noshow_threshold_1",
  "noshow_threshold_2",
  "noshow_low_1",
  "noshow_low_2",
  "noshow_high_1",
  "noshow_high_2"
)

missing_background_columns <- setdiff(
  required_background_columns,
  names(seed_outcomes)
)

if (length(missing_background_columns) > 0) {
  stop(
    "The following required background columns are missing: ",
    paste(missing_background_columns, collapse = ", ")
  )
}

policy_labels <- c(
  baseline = "Baseline",
  horizon_only = "Horizon only",
  reservation_only = "Reservation only",
  both_flexible = "Both flexible"
)

policy_order <- unname(policy_labels)

active_comparison_to_policy <- c(
  horizon_only_vs_baseline = "Horizon only",
  reservation_only_vs_baseline = "Reservation only",
  both_flexible_vs_baseline = "Both flexible"
)

pct <- function(x, digits = 1) {
  paste0(formatC(100 * x, format = "f", digits = digits), "%")
}

pp <- function(x, digits = 2) {
  paste0(formatC(100 * x, format = "f", digits = digits), " pp")
}

theme_report <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot",
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      axis.title = element_text(face = "bold")
    )
}

add_characteristic_flags <- function(data) {
  data |>
    mutate(
      no_show_sensitive =
        noshow_threshold_1 < noshow_threshold_2 &
        noshow_low_1 >= noshow_low_2 &
        noshow_high_1 >= noshow_high_2,

      no_show_sensitive_strict =
        noshow_threshold_1 < noshow_threshold_2 &
        noshow_low_1 > noshow_low_2 &
        noshow_high_1 > noshow_high_2,

      balking_sensitive =
        balk_threshold_1 < balk_threshold_2 &
        balk_low_1 >= balk_low_2 &
        balk_high_1 >= balk_high_2,

      balking_sensitive_strict =
        balk_threshold_1 < balk_threshold_2 &
        balk_low_1 > balk_low_2 &
        balk_high_1 > balk_high_2,

      cancellation_prone =
        cancel_1 > cancel_2,

      jointly_delay_sensitive =
        no_show_sensitive & balking_sensitive,

      jointly_delay_sensitive_strict =
        no_show_sensitive_strict & balking_sensitive_strict
    )
}

make_characteristic_rows <- function(data) {
  flagged <- add_characteristic_flags(data)

  bind_rows(
    flagged |>
      filter(no_show_sensitive) |>
      mutate(
        characteristic = "Class 1 more no-show sensitive",
        characteristic_level = "Main definition"
      ),

    flagged |>
      filter(balking_sensitive) |>
      mutate(
        characteristic = "Class 1 more balking sensitive",
        characteristic_level = "Main definition"
      ),

    flagged |>
      filter(cancellation_prone) |>
      mutate(
        characteristic = "Class 1 more cancellation prone",
        characteristic_level = "Class 1 cancellation > Class 2"
      ),

    flagged |>
      mutate(
        characteristic = "Priority-group share",
        characteristic_level = percent(class1_share, accuracy = 1)
      ),

    flagged |>
      filter(jointly_delay_sensitive) |>
      mutate(
        characteristic = "Class 1 jointly delay sensitive",
        characteristic_level = "No-show and balking sensitive"
      )
  ) |>
    mutate(
      characteristic = factor(
        characteristic,
        levels = c(
          "Class 1 more no-show sensitive",
          "Class 1 more balking sensitive",
          "Class 1 more cancellation prone",
          "Priority-group share",
          "Class 1 jointly delay sensitive"
        )
      )
    )
}

average_seed_outcomes <- seed_outcomes |>
  filter(selection_objective == "average_utilization") |>
  mutate(
    policy_label = factor(
      policy_labels[policy],
      levels = policy_order
    ),
    average_utilization_value = .data[[average_metric_col]],
    class1_served_rate_value = .data[[class1_served_col]],
    class2_served_rate_value = .data[[class2_served_col]]
  )

baseline_seed <- average_seed_outcomes |>
  filter(policy == "baseline") |>
  select(
    background_id,
    all_of(seed_col),
    baseline_average_utilization = average_utilization_value,
    baseline_class1_served_rate = class1_served_rate_value,
    baseline_class2_served_rate = class2_served_rate_value
  )

average_seed_outcomes <- average_seed_outcomes |>
  left_join(
    baseline_seed,
    by = c("background_id", seed_col)
  ) |>
  mutate(
    utilization_gain =
      average_utilization_value - baseline_average_utilization,
    class1_served_rate_gain =
      class1_served_rate_value - baseline_class1_served_rate,
    class2_served_rate_gain =
      class2_served_rate_value - baseline_class2_served_rate
  )

characteristic_seed_outcomes <- make_characteristic_rows(
  average_seed_outcomes
)

background_policy_means <- average_seed_outcomes |>
  group_by(background_id, policy, policy_label) |>
  summarise(
    average_utilization =
      mean(average_utilization_value, na.rm = TRUE),
    utilization_gain =
      mean(utilization_gain, na.rm = TRUE),
    class1_served_rate_gain =
      mean(class1_served_rate_gain, na.rm = TRUE),
    class2_served_rate_gain =
      mean(class2_served_rate_gain, na.rm = TRUE),
    .groups = "drop"
  ) |>
  group_by(background_id) |>
  mutate(
    best_average_utilization =
      max(average_utilization, na.rm = TRUE),
    practically_best =
      average_utilization >= best_average_utilization - 0.005
  ) |>
  ungroup()

background_characteristics <- average_seed_outcomes |>
  filter(policy == "baseline") |>
  distinct(
    background_id,
    across(all_of(required_background_columns[-1]))
  ) |>
  make_characteristic_rows() |>
  select(
    background_id,
    characteristic,
    characteristic_level,
    rho,
    class1_share
  )

characteristic_policy_means <- background_policy_means |>
  inner_join(
    background_characteristics,
    by = "background_id"
  )

average_policy_outcomes <- policy_outcomes |>
  filter(selection_objective == "average_utilization") |>
  mutate(
    policy_label = factor(
      policy_labels[policy],
      levels = policy_order
    )
  ) |>
  make_characteristic_rows()

average_pairwise <- pairwise |>
  filter(
    selection_objective == "average_utilization",
    comparison %in% names(active_comparison_to_policy)
  ) |>
  mutate(
    policy_label = factor(
      active_comparison_to_policy[comparison],
      levels = policy_order[-1]
    )
  ) |>
  make_characteristic_rows()
```

## Research question

Under which patient-population conditions do the following policies improve average appointment utilization?

| Policy | Booking horizon | Same-day-release reservations |
|---|---|---|
| Baseline | Native horizon | None |
| Horizon only | Optimized | None |
| Reservation only | Native horizon | Quantity and window optimized |
| Both flexible | Horizon, quantity, and window jointly optimized |

The main outcome is:

> **Average utilization** = **completed appointments** ÷ **available measured appointment slots**

A difference smaller than **0.005 average utilization** is treated as practically equivalent. Policies that fall within 0.005 of the best result for a background are therefore reported as tied for best rather than forced into a single ranking.

## Patient characteristics

This Phase 1 analysis uses the existing heterogeneous simulation bank. It identifies backgrounds matching five clinically interpretable Class 1 characteristics:

```{r}
#| label: tbl-characteristic-definitions

characteristic_definitions <- tibble(
  `Patient characteristic` = c(
    "Class 1 more no-show sensitive",
    "Class 1 more balking sensitive",
    "Class 1 more cancellation prone",
    "Priority-group share",
    "Class 1 jointly delay sensitive"
  ),
  `Phase 1 definition` = c(
    paste(
      "Lower no-show threshold, with pre- and post-threshold",
      "no-show probabilities at least as high as Class 2"
    ),
    paste(
      "Lower balking threshold, with pre- and post-threshold",
      "balking probabilities at least as high as Class 2"
    ),
    "Class 1 cancellation probability is higher than Class 2",
    "Class 1 share analysed at 0.1, 0.3, 0.5, 0.7 and 0.9",
    paste(
      "Class 1 satisfies both the no-show-sensitive and",
      "balking-sensitive definitions"
    )
  ),
  `Policy interaction being tested` = c(
    "Whether shorter horizons, near-term reservations, or both reduce delay-related no-shows",
    "Whether the policies make acceptable near-term appointments more available",
    "Whether day-of release prevents unused Class 1 reservations from lowering utilization",
    "Whether the value and appropriate size of reservations depend on the size of Class 1",
    "Whether combining horizon flexibility and reservations is especially valuable when both delay responses are elevated"
  )
)

kable(
  characteristic_definitions,
  align = c("l", "l", "l")
)
```

These are exploratory associations. In the existing bank, non-target patient behaviours were not held equal between the classes. The later targeted simulations will use controlled same, mild and strong Class 1 contrasts.

## Simulation coverage

```{r}
#| label: tbl-profile-coverage

profile_coverage <- background_characteristics |>
  count(
    characteristic,
    characteristic_level,
    name = "Backgrounds"
  ) |>
  arrange(characteristic, characteristic_level)

kable(
  profile_coverage,
  col.names = c(
    "Patient characteristic",
    "Characteristic level",
    "Backgrounds"
  ),
  align = c("l", "l", "r")
)
```

## All four policies side by side

The first figure keeps all four policies in one comparison. Each point is the median within-background utilization difference from baseline; thick intervals show the interquartile range across backgrounds.

```{r}
#| label: fig-four-policy-characteristics
#| fig-cap: "All four policies compared within each patient characteristic. Baseline is zero by construction. Points show median utilization differences from baseline; intervals show the interquartile range across backgrounds."
#| fig-width: 12
#| fig-height: 9

four_policy_summary <- characteristic_policy_means |>
  group_by(
    characteristic,
    characteristic_level,
    policy_label
  ) |>
  summarise(
    n_backgrounds = n_distinct(background_id),
    q25_pp = 100 * quantile(
      utilization_gain,
      0.25,
      na.rm = TRUE
    ),
    median_pp = 100 * median(
      utilization_gain,
      na.rm = TRUE
    ),
    q75_pp = 100 * quantile(
      utilization_gain,
      0.75,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

ggplot(
  four_policy_summary,
  aes(
    x = policy_label,
    y = median_pp,
    ymin = q25_pp,
    ymax = q75_pp,
    group = characteristic_level,
    shape = characteristic_level
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    colour = "grey70"
  ) +
  geom_pointrange(
    position = position_dodge(width = 0.55),
    linewidth = 0.55
  ) +
  facet_wrap(
    ~characteristic,
    scales = "free_y",
    ncol = 2
  ) +
  labs(
    x = NULL,
    y = "Average-utilization difference from baseline (pp)",
    shape = "Characteristic level"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

This is a descriptive comparison of the four selected policies. Confidence-supported claims between specific policies can still use the paired bootstrap results, but the report does not organise the presentation as six separate pairwise analyses.

## Which policies are practically tied for best?

A policy is counted as practically best when its average utilization is no more than 0.005 below the highest-utilization policy in that background. Multiple policies can be counted in the same background.

```{r}
#| label: fig-practically-best
#| fig-cap: "Share of backgrounds in which each policy is within 0.005 average utilization of the best policy. Percentages may sum to more than 100% because practical ties are retained."
#| fig-width: 12
#| fig-height: 9

practically_best_summary <- characteristic_policy_means |>
  group_by(
    characteristic,
    characteristic_level,
    policy_label
  ) |>
  summarise(
    share_practically_best =
      mean(practically_best, na.rm = TRUE),
    n_backgrounds =
      n_distinct(background_id),
    .groups = "drop"
  )

ggplot(
  practically_best_summary,
  aes(
    x = policy_label,
    y = share_practically_best,
    group = characteristic_level,
    fill = characteristic_level
  )
) +
  geom_col(
    position = position_dodge(width = 0.75),
    width = 0.68
  ) +
  facet_wrap(
    ~characteristic,
    ncol = 2
  ) +
  scale_y_continuous(
    labels = percent_format(),
    limits = c(0, 1)
  ) +
  labs(
    x = NULL,
    y = "Share practically tied for best",
    fill = "Characteristic level"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

## How demand and priority-group share change the comparison

The following heatmap presents all four policies together. Baseline remains the zero reference. Values are grouped medians over the remaining heterogeneous parameters.

```{r}
#| label: fig-policy-demand-share
#| fig-cap: "Median average-utilization difference from baseline by demand-to-capacity ratio, priority-group share and policy. Each cell aggregates over the remaining scenario parameters."
#| fig-width: 13
#| fig-height: 10

demand_share_summary <- background_policy_means |>
  inner_join(
    average_seed_outcomes |>
      filter(policy == "baseline") |>
      distinct(
        background_id,
        rho,
        class1_share
      ),
    by = "background_id"
  ) |>
  group_by(
    policy_label,
    rho,
    class1_share
  ) |>
  summarise(
    median_gain_pp =
      100 * median(utilization_gain, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    rho = factor(rho),
    class1_share = factor(class1_share)
  )

heat_limit <- quantile(
  abs(demand_share_summary$median_gain_pp),
  0.95,
  na.rm = TRUE
)

ggplot(
  demand_share_summary,
  aes(
    x = class1_share,
    y = rho,
    fill = median_gain_pp
  )
) +
  geom_tile(
    colour = "white",
    linewidth = 0.25
  ) +
  geom_text(
    aes(
      label = formatC(
        median_gain_pp,
        format = "f",
        digits = 1
      )
    ),
    size = 2.8
  ) +
  facet_wrap(
    ~policy_label,
    ncol = 2
  ) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "white",
    high = "#2166AC",
    midpoint = 0,
    limits = c(-heat_limit, heat_limit),
    oob = squish,
    name = "Median\ndifference (pp)"
  ) +
  labs(
    x = "Priority-group share",
    y = "Demand-to-capacity ratio"
  ) +
  theme_report()
```

## Results by patient characteristic

### Class 1 more no-show sensitive

```{r}
#| label: fig-noshow-sensitive
#| fig-cap: "Four-policy comparison when Class 1 has a lower no-show threshold and no-show probabilities at least as high as Class 2."
#| fig-width: 11
#| fig-height: 5.5

plot_characteristic_by_demand <- function(
    characteristic_name,
    caption_y = "Median utilization difference from baseline (pp)") {

  plot_data <- characteristic_policy_means |>
    filter(characteristic == characteristic_name) |>
    group_by(
      characteristic_level,
      rho,
      policy_label
    ) |>
    summarise(
      median_gain_pp =
        100 * median(utilization_gain, na.rm = TRUE),
      q25_pp =
        100 * quantile(utilization_gain, 0.25, na.rm = TRUE),
      q75_pp =
        100 * quantile(utilization_gain, 0.75, na.rm = TRUE),
      .groups = "drop"
    )

  ggplot(
    plot_data,
    aes(
      x = factor(rho),
      y = median_gain_pp,
      group = policy_label,
      shape = policy_label
    )
  ) +
    geom_hline(yintercept = 0, colour = "grey55") +
    geom_hline(
      yintercept = 0.5,
      linetype = "dashed",
      colour = "grey70"
    ) +
    geom_line(linewidth = 0.7) +
    geom_point(size = 2) +
    facet_wrap(
      ~characteristic_level,
      scales = "free_y"
    ) +
    labs(
      x = "Demand-to-capacity ratio",
      y = caption_y,
      shape = "Policy"
    ) +
    theme_report()
}

plot_characteristic_by_demand(
  "Class 1 more no-show sensitive"
)
```

### Class 1 more balking sensitive

```{r}
#| label: fig-balking-sensitive
#| fig-cap: "Four-policy comparison when Class 1 has a lower balking threshold and balking probabilities at least as high as Class 2."
#| fig-width: 11
#| fig-height: 5.5

plot_characteristic_by_demand(
  "Class 1 more balking sensitive"
)
```

### Class 1 more cancellation prone

```{r}
#| label: fig-cancellation-prone
#| fig-cap: "Four-policy comparison when Class 1 has a higher cancellation probability than Class 2."
#| fig-width: 11
#| fig-height: 5.5

plot_characteristic_by_demand(
  "Class 1 more cancellation prone"
)
```

### Priority-group share

The share analysis uses the same four-policy outcomes but treats each tested Class 1 share as a separate patient-composition level.

```{r}
#| label: fig-priority-share
#| fig-cap: "Four-policy comparison by Class 1 share and demand-to-capacity ratio."
#| fig-width: 12
#| fig-height: 7

share_plot_data <- characteristic_policy_means |>
  filter(characteristic == "Priority-group share") |>
  group_by(
    characteristic_level,
    rho,
    policy_label
  ) |>
  summarise(
    median_gain_pp =
      100 * median(utilization_gain, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  share_plot_data,
  aes(
    x = factor(rho),
    y = median_gain_pp,
    group = policy_label,
    shape = policy_label
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    colour = "grey70"
  ) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.8) +
  facet_wrap(
    ~characteristic_level,
    ncol = 3,
    scales = "free_y"
  ) +
  labs(
    x = "Demand-to-capacity ratio",
    y = "Median utilization difference from baseline (pp)",
    shape = "Policy"
  ) +
  theme_report()
```

### Class 1 jointly delay sensitive

```{r}
#| label: fig-jointly-sensitive
#| fig-cap: "Four-policy comparison when Class 1 is both more no-show sensitive and more balking sensitive than Class 2."
#| fig-width: 11
#| fig-height: 5.5

plot_characteristic_by_demand(
  "Class 1 jointly delay sensitive"
)
```

## Effects on Class 1 and Class 2

The main policy presentation compares all four policies together. Patient-group confidence classifications remain relative to baseline because a gain or loss must be defined against a reference policy.

```{r}
#| label: fig-group-effects-characteristics
#| fig-cap: "Confidence-supported Class 1 and Class 2 served-rate patterns for each active policy compared with baseline, shown by patient characteristic."
#| fig-width: 13
#| fig-height: 10

collapse_group_effect <- function(x) {
  case_when(
    x == "both_groups_gain" ~
      "Both groups gain",

    x == "priority_gains_general_loses" ~
      "Class 1 gains; Class 2 loses",

    x == "priority_loses_general_gains" ~
      "Class 2 gains; Class 1 loses",

    x %in% c(
      "priority_gains_general_tied",
      "priority_tied_general_gains"
    ) ~
      "One group gains; other ties",

    x %in% c(
      "priority_loses_general_tied",
      "priority_tied_general_loses",
      "both_groups_lose"
    ) ~
      "Loss with no supported gain",

    TRUE ~
      "No clear group change"
  )
}

group_effect_levels <- c(
  "Both groups gain",
  "Class 1 gains; Class 2 loses",
  "Class 2 gains; Class 1 loses",
  "One group gains; other ties",
  "Loss with no supported gain",
  "No clear group change"
)

group_effect_summary <- average_pairwise |>
  mutate(
    group_effect = factor(
      collapse_group_effect(
        patient_group_supported_tradeoff
      ),
      levels = group_effect_levels
    )
  ) |>
  count(
    characteristic,
    characteristic_level,
    policy_label,
    group_effect,
    name = "n"
  ) |>
  group_by(
    characteristic,
    characteristic_level,
    policy_label
  ) |>
  mutate(
    share = n / sum(n)
  ) |>
  ungroup()

ggplot(
  group_effect_summary |>
    filter(characteristic != "Priority-group share"),
  aes(
    x = policy_label,
    y = share,
    fill = group_effect
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic,
    ncol = 2
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of matching backgrounds",
    fill = "Supported group pattern"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

“Loss with no supported gain” combines:

- Class 1 loses while Class 2 ties;
- Class 2 loses while Class 1 ties; and
- both groups lose.

It excludes the two explicit gain-loss trade-off categories.

```{r}
#| label: fig-group-effects-share
#| fig-cap: "Confidence-supported Class 1 and Class 2 served-rate patterns by priority-group share."
#| fig-width: 13
#| fig-height: 8

ggplot(
  group_effect_summary |>
    filter(characteristic == "Priority-group share"),
  aes(
    x = policy_label,
    y = share,
    fill = group_effect
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic_level,
    ncol = 3
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of backgrounds",
    fill = "Supported group pattern"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

## Selected policy design

\(Q\) is the number of slots per service day reserved for Class 1 during the reservation window. Under same-day release, unused Class 1 reservations return to the pooled appointment capacity on the service day.

The booking horizon was selected from the discrete candidate grid **2, 6, 10, 14, 18, 22 and 26 days**. Intermediate values were not tested.

### Booking-horizon selections

```{r}
#| label: fig-selected-horizons
#| fig-cap: "Distribution of selected booking horizons by policy and patient characteristic. The horizon is categorical because only seven values were tested."
#| fig-width: 13
#| fig-height: 10

horizon_distribution <- average_policy_outcomes |>
  filter(characteristic != "Priority-group share") |>
  count(
    characteristic,
    policy_label,
    selected_horizon_days,
    name = "n"
  ) |>
  group_by(
    characteristic,
    policy_label
  ) |>
  mutate(
    share = n / sum(n)
  ) |>
  ungroup() |>
  mutate(
    selected_horizon_days =
      factor(
        selected_horizon_days,
        levels = c(2, 6, 10, 14, 18, 22, 26)
      )
  )

ggplot(
  horizon_distribution,
  aes(
    x = policy_label,
    y = share,
    fill = selected_horizon_days
  )
) +
  geom_col(width = 0.72) +
  facet_wrap(
    ~characteristic,
    ncol = 2
  ) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    x = NULL,
    y = "Share of matching backgrounds",
    fill = "Selected horizon\n(days)"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

### Reservation quantity and window

```{r}
#| label: fig-reservation-design
#| fig-cap: "Selected reservation quantity and reservation-window length for reservation-only and both-flexible policies. Points show medians and intervals show interquartile ranges."
#| fig-width: 13
#| fig-height: 9

reservation_design <- bind_rows(
  average_policy_outcomes |>
    filter(
      characteristic != "Priority-group share",
      policy %in% c(
        "reservation_only",
        "both_flexible"
      )
    ) |>
    mutate(
      metric = "Reserved quantity (% of daily capacity)",
      value = 100 * selected_Q / slots_per_day
    ),

  average_policy_outcomes |>
    filter(
      characteristic != "Priority-group share",
      policy %in% c(
        "reservation_only",
        "both_flexible"
      ),
      selected_Q > 0,
      selected_window >= 1
    ) |>
    mutate(
      metric = "Reservation-window length (days)",
      value = selected_window
    )
) |>
  group_by(
    characteristic,
    policy_label,
    metric
  ) |>
  summarise(
    q25 = quantile(value, 0.25, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    q75 = quantile(value, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  reservation_design,
  aes(
    x = policy_label,
    y = median,
    ymin = q25,
    ymax = q75
  )
) +
  geom_pointrange(linewidth = 0.6) +
  facet_grid(
    metric ~ characteristic,
    scales = "free_y"
  ) +
  labs(
    x = NULL,
    y = "Selected parameter"
  ) +
  theme_report() +
  theme(
    axis.text.x = element_text(
      angle = 25,
      hjust = 1
    )
  )
```

## Strict-definition sensitivity check

The main no-show and balking definitions allow one or both probabilities to tie with Class 2. The strict definitions require both probabilities to be higher.

```{r}
#| label: tbl-strict-definition-coverage

strict_coverage <- average_seed_outcomes |>
  filter(policy == "baseline") |>
  distinct(
    background_id,
    across(all_of(required_background_columns[-1]))
  ) |>
  add_characteristic_flags() |>
  summarise(
    `No-show sensitive: main` =
      sum(no_show_sensitive),
    `No-show sensitive: strict` =
      sum(no_show_sensitive_strict),
    `Balking sensitive: main` =
      sum(balking_sensitive),
    `Balking sensitive: strict` =
      sum(balking_sensitive_strict),
    `Jointly sensitive: main` =
      sum(jointly_delay_sensitive),
    `Jointly sensitive: strict` =
      sum(jointly_delay_sensitive_strict)
  ) |>
  pivot_longer(
    everything(),
    names_to = "Definition",
    values_to = "Backgrounds"
  )

kable(
  strict_coverage,
  align = c("l", "r")
)
```

The strict subsets should be used as a robustness check rather than as the main analysis if their sample sizes are too small for stable scenario-level summaries.

## Interpretation for the next simulation stage

This document should be used to identify:

- which characteristics show the clearest separation among the four policies;
- where policies are practically tied;
- which demand ratios and Class 1 shares should receive denser coverage;
- whether the selected horizon, \(Q\), and reservation window respond consistently to the patient characteristic; and
- where the Class 1 benefit is accompanied by a Class 2 loss.

The next targeted simulation bank should then hold non-target patient behaviour equal within a background and vary the target Class 1 characteristic through **same, mild and strong** contrasts relative to two Class 2 reference profiles.
