---
title: "H1 Policy Comparison: Strict Priority Reservations"
subtitle: "Policies optimized for average appointment utilization"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show analysis code"
    embed-resources: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
  freeze: auto
---


```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(scales)
library(knitr)

find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = FALSE)

  repeat {
    marker <- file.path(
      current,
      "full_run_summaries",
      "h1_policy_outcomes"
    )

    if (dir.exists(marker)) {
      return(current)
    }

    parent <- dirname(current)

    if (identical(parent, current)) {
      stop(
        "Could not find the repository root. Render this file from inside ",
        "CUIMC-Appointment-Simulation."
      )
    }

    current <- parent
  }
}

repo_root <- find_repo_root()
data_root <- file.path(
  repo_root,
  "full_run_summaries",
  "h1_policy_outcomes"
)

variant <- "strict"
companion_variant <- "release"

policy_outcomes <- read_csv(
  file.path(data_root, variant, "selected_policy_outcomes.csv"),
  show_col_types = FALSE
)

pairwise <- read_csv(
  file.path(data_root, variant, "pairwise_group_deltas.csv"),
  show_col_types = FALSE
)

objective_switch <- read_csv(
  file.path(data_root, variant, "objective_switch_deltas.csv"),
  show_col_types = FALSE
)

companion_outcomes <- read_csv(
  file.path(data_root, companion_variant, "selected_policy_outcomes.csv"),
  show_col_types = FALSE
)

companion_pairwise <- read_csv(
  file.path(data_root, companion_variant, "pairwise_group_deltas.csv"),
  show_col_types = FALSE
)

companion_switch <- read_csv(
  file.path(data_root, companion_variant, "objective_switch_deltas.csv"),
  show_col_types = FALSE
)

avg_outcomes <- policy_outcomes |>
  filter(selection_objective == "average_utilization")

avg_pairwise <- pairwise |>
  filter(selection_objective == "average_utilization")

avg_companion_pairwise <- companion_pairwise |>
  filter(selection_objective == "average_utilization")

policy_labels <- c(
  baseline = "Baseline",
  horizon_only = "Horizon only",
  reservation_only = "Reservation only",
  both_flexible = "Both flexible"
)

comparison_labels <- c(
  horizon_only_vs_baseline = "Horizon only − baseline",
  reservation_only_vs_baseline = "Reservation only − baseline",
  both_flexible_vs_baseline = "Both flexible − baseline",
  both_flexible_vs_horizon_only = "Both flexible − horizon only",
  both_flexible_vs_reservation_only = "Both flexible − reservation only",
  reservation_only_vs_horizon_only = "Reservation only − horizon only"
)

pct <- function(x, digits = 1) {
  paste0(formatC(100 * x, format = "f", digits = digits), "%")
}

pp <- function(x, digits = 2) {
  paste0(formatC(100 * x, format = "f", digits = digits), " pp")
}

comparison_rows <- function(comparison_name) {
  avg_pairwise |>
    filter(comparison == comparison_name)
}

supported_rate <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = mean(
        average_utilization_change_status == "supported_increase"
      )
    ) |>
    pull(value)
}

decrease_rate <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = mean(
        average_utilization_change_status == "supported_decrease"
      )
    ) |>
    pull(value)
}

tie_rate <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = mean(
        average_utilization_change_status == "practical_tie"
      )
    ) |>
    pull(value)
}

median_gain <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(value = median(delta_average_utilization, na.rm = TRUE)) |>
    pull(value)
}

supported_count <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = sum(
        average_utilization_change_status == "supported_increase"
      )
    ) |>
    pull(value)
}

decrease_count <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = sum(
        average_utilization_change_status == "supported_decrease"
      )
    ) |>
    pull(value)
}

tie_count <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = sum(
        average_utilization_change_status == "practical_tie"
      )
    ) |>
    pull(value)
}

tradeoff_rate <- function(comparison_name) {
  comparison_rows(comparison_name) |>
    summarise(
      value = mean(
        patient_group_supported_tradeoff ==
          "priority_gains_general_loses"
      )
    ) |>
    pull(value)
}

switch_policy_rows <- function(policy_name) {
  objective_switch |>
    filter(policy == policy_name)
}

same_cell_rate <- function(policy_name) {
  switch_policy_rows(policy_name) |>
    summarise(value = mean(same_policy_cell, na.rm = TRUE)) |>
    pull(value)
}

median_switch_average <- function(policy_name) {
  switch_policy_rows(policy_name) |>
    summarise(value = median(delta_average_utilization, na.rm = TRUE)) |>
    pull(value)
}

median_switch_weighted <- function(policy_name) {
  switch_policy_rows(policy_name) |>
    summarise(value = median(delta_weighted_utilization, na.rm = TRUE)) |>
    pull(value)
}

theme_report <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot",
      legend.position = "bottom",
      strip.text = element_text(face = "bold"),
      axis.title = element_text(face = "bold")
    )
}
```

## Main findings

- **Most policy comparisons are ties.** Both-flexible improves on baseline in `r pct(supported_rate("both_flexible_vs_baseline"))` of backgrounds and is practically tied in `r pct(tie_rate("both_flexible_vs_baseline"))`.
- **Large gains are concentrated under heavy demand.** The median both-flexible gain is `r pp(median_gain("both_flexible_vs_baseline"))` overall, but the upper-demand scenarios contain the large improvements.
- **Neither single policy uniformly dominates.** Reservation-only beats horizon-only in `r supported_count("reservation_only_vs_horizon_only")` backgrounds, horizon-only beats reservation-only in `r decrease_count("reservation_only_vs_horizon_only")`, and they tie in `r tie_count("reservation_only_vs_horizon_only")`.
- **Reservations often redistribute access.** Priority served rate rises while general-group served rate falls in `r pct(tradeoff_rate("reservation_only_vs_baseline"))` of reservation-only comparisons with baseline.
- **The twice-weighted Class 1 objective is materially different.** Exact policy cells match only `r pct(same_cell_rate("reservation_only"))` of the time for reservation-only and `r pct(same_cell_rate("both_flexible"))` for both-flexible.

## How the simulations were run

- **840 backgrounds:** 42 anchor settings and 798 Sobol-sampled settings.
- A **policy cell** is one exact combination of booking horizon, reserved quantity, and reservation-window length.
- Every evaluated policy cell was run with **10 paired seeds**, so competing policies were compared under the same random seeds.
- The final postprocessed file contains **67,200 selected-policy seed records per reservation rule**: 840 backgrounds × 4 policies × 2 optimization objectives × 10 seeds. These are the selected outcomes; the optimizer ran additional candidate cells before choosing each optimum.
- Demand-to-capacity ratio ranged from **0.8 to 3.0**; priority-group share ranged from **0.1 to 0.9**.
- Daily capacity ranged from **20 to 50 slots** and the native horizon ranged from **2 to 26 days**.
- Cancellation, balking, and no-show behavior varied across both groups.
- Paired bootstrap intervals used **2,000 draws**. Bootstrap draws are resamples of the paired seed differences, not additional simulation runs. A difference within **0.005** utilization was treated as a practical tie.


### Problem parameters

The 840 backgrounds were sampled from the following parameter grid. The design was not a full factorial: it combined 42 anchor scenarios with 798 Sobol-sampled scenarios.

```{r}
#| label: tbl-problem-parameters

problem_parameters <- tibble(
  Parameter = c(
    "Demand-to-capacity ratio (ρ)",
    "Native booking horizon (days)",
    "Priority-group share (Class 1)",
    "Daily capacity (slots/day)",
    "Cancellation probability (per class)",
    "Balking threshold (per class, days)",
    "Pre-threshold balking probability (per class)",
    "Post-threshold balking probability (per class)",
    "No-show threshold (per class, days)",
    "Pre-threshold no-show probability (per class)",
    "Post-threshold no-show probability (per class)"
  ),
  `Tested values` = c(
    "0.8, 1.0, 1.2, 1.4, 1.6, 2.0, 3.0",
    "2, 6, 10, 14, 18, 22, 26",
    "0.1, 0.3, 0.5, 0.7, 0.9",
    "20, 30, 40, 50",
    "0.1, 0.2, 0.3",
    "5, 9, 16, 22, 24",
    "0.05, 0.1, 0.2, 0.3",
    "0.1, 0.2, 0.3, 0.4",
    "4, 7, 14, 20, 22",
    "0.05, 0.1, 0.2, 0.3",
    "0.1, 0.2, 0.3, 0.4"
  )
)

kable(
  problem_parameters,
  align = c("l", "l"),
  col.names = c("Parameter", "Tested values")
)
```

The cancellation, balking, and no-show parameters were sampled separately for Class 1 and Class 2. Valid sampled backgrounds required each post-threshold probability to exceed its corresponding pre-threshold probability and each class's balking threshold to exceed its no-show threshold. When a sampled threshold was outside a tested booking horizon, the simulation capped it at the active horizon minus one day.

### Policy-search parameters

The optimized policy search used:

```{r}
#| label: tbl-policy-search-parameters

policy_search_parameters <- tibble(
  `Policy parameter` = c(
    "Booking horizon",
    "Reserved quantity Q",
    "Reservation-window length"
  ),
  `Candidate values` = c(
    "2, 6, 10, 14, 18, 22, 26 days",
    "0 to daily capacity; positive Q searched coarsely in steps of 5 and refined in steps of 1",
    "1 day through the selected horizon; searched coarsely in steps of 2 and refined in steps of 1"
  )
)

kable(
  policy_search_parameters,
  align = c("l", "l")
)
```

The four policy regimes were:

| Policy | Booking horizon | Priority reservations |
|---|---|---|
| Baseline | Fixed at the background value | None |
| Horizon only | Optimized | None |
| Reservation only | Fixed at the background value | Quantity and reservation window optimized |
| Both flexible | Optimized | Quantity and reservation window optimized |

The primary objective was **average utilization**:

> **Average utilization** = **completed appointments** ÷ **available measured appointment slots**

Class 1 and Class 2 remain different simulated patient populations in every regime. Class 1 is called the **priority group** only when `Q > 0`. When `Q = 0`, both groups book through the general first-come-first-served pool.

## Overall policy comparison

```{r}
#| label: fig-overall-comparisons
#| fig-cap: "Distribution of average-utilization differences across all six policy comparisons. Points show medians, thick lines show interquartile ranges, and thin lines show the 10th–90th percentiles."
#| fig-width: 10
#| fig-height: 6

comparison_order <- unname(comparison_labels)

comparison_summary <- avg_pairwise |>
  group_by(comparison) |>
  summarise(
    q10 = quantile(delta_average_utilization, 0.10, na.rm = TRUE),
    q25 = quantile(delta_average_utilization, 0.25, na.rm = TRUE),
    median = median(delta_average_utilization, na.rm = TRUE),
    q75 = quantile(delta_average_utilization, 0.75, na.rm = TRUE),
    q90 = quantile(delta_average_utilization, 0.90, na.rm = TRUE),
    supported = mean(
      average_utilization_change_status == "supported_increase"
    ),
    practical_tie = mean(
      average_utilization_change_status == "practical_tie"
    ),
    .groups = "drop"
  ) |>
  mutate(
    comparison_label = comparison_labels[comparison],
    comparison_label = factor(
      comparison_label,
      levels = rev(comparison_order)
    ),
    comparison_group = if_else(
      grepl("_vs_baseline$", comparison),
      "Compared with baseline",
      "Compared with another active policy"
    ),
    across(c(q10, q25, median, q75, q90), ~100 * .x),
    result_label = paste0(
      pct(supported),
      " supported; ",
      pct(practical_tie),
      " tied"
    )
  )

ggplot(
  comparison_summary,
  aes(y = comparison_label)
) +
  geom_vline(xintercept = 0, colour = "grey45") +
  geom_vline(
    xintercept = 0.5,
    linetype = "dashed",
    colour = "grey65"
  ) +
  geom_segment(
    aes(x = q10, xend = q90, yend = comparison_label),
    linewidth = 0.6,
    colour = "grey45"
  ) +
  geom_segment(
    aes(x = q25, xend = q75, yend = comparison_label),
    linewidth = 3
  ) +
  geom_point(aes(x = median), size = 2.5) +
  geom_text(
    aes(x = q90, label = result_label),
    hjust = -0.08,
    size = 3.1
  ) +
  facet_wrap(
    ~comparison_group,
    ncol = 1,
    scales = "free_y"
  ) +
  scale_x_continuous(
    name = "Average-utilization difference (percentage points)",
    expand = expansion(mult = c(0.05, 0.34))
  ) +
  labs(y = NULL) +
  coord_cartesian(clip = "off") +
  theme_report()
```

- Both-flexible weakly contains the other policy choices, so it cannot be worse after optimization.
- That does **not** mean both policies are always needed. Most backgrounds are practical ties.
- Reservation-only and horizon-only should be read as competing single-policy options, not as a ranking that applies to every clinic.

```{r}
#| label: tbl-comparison-summary

comparison_table <- comparison_summary |>
  transmute(
    Comparison = as.character(comparison_label),
    `Median difference` = paste0(
      formatC(median, format = "f", digits = 2),
      " pp"
    ),
    `Supported increase` = pct(supported),
    `Practical tie` = pct(practical_tie)
  )

kable(
  comparison_table,
  align = c("l", "r", "r", "r")
)
```

## Which scenarios benefit?

```{r}
#| label: fig-scenario-heatmap
#| fig-cap: "Median average-utilization difference by demand pressure and priority-group share. Values are aggregated over the remaining scenario parameters."
#| fig-width: 11
#| fig-height: 8

heatmap_data <- avg_pairwise |>
  group_by(comparison, rho, class1_share) |>
  summarise(
    median_delta_pp =
      100 * median(delta_average_utilization, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    comparison_label = comparison_labels[comparison],
    class1_share = factor(class1_share),
    rho = factor(rho)
  )

heat_limit <- quantile(
  abs(heatmap_data$median_delta_pp),
  0.95,
  na.rm = TRUE
)

ggplot(
  heatmap_data,
  aes(
    x = class1_share,
    y = rho,
    fill = median_delta_pp
  )
) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_text(
    aes(
      label = if_else(
        abs(median_delta_pp) >= 0.5,
        formatC(
          median_delta_pp,
          format = "f",
          digits = 1
        ),
        ""
      )
    ),
    size = 2.7
  ) +
  facet_wrap(
    ~comparison_label,
    ncol = 3
  ) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "white",
    high = "#2166AC",
    midpoint = 0,
    limits = c(-heat_limit, heat_limit),
    oob = squish,
    name = "Median\nchange (pp)"
  ) +
  labs(
    x = "Priority-group share",
    y = "Demand-to-capacity ratio"
  ) +
  theme_report()
```

The dominant pattern is demand pressure:

- At low and moderate demand, the policies are usually tied in utilization.
- The clearest gains appear when demand approaches the upper end of the tested range.
- Priority-group share changes which single policy performs better, but it does not create one universal winner.

### No-show behavior and reservation value

```{r}
#| label: fig-noshow-pattern
#| fig-cap: "Reservation-related gains rise under heavy demand when the general group has a higher low-delay no-show probability. These are grouped associations across backgrounds, not isolated one-parameter causal effects."
#| fig-width: 10
#| fig-height: 4.8

noshow_pattern <- avg_pairwise |>
  filter(
    comparison %in% c(
      "reservation_only_vs_baseline",
      "both_flexible_vs_baseline"
    )
  ) |>
  group_by(
    comparison,
    rho,
    noshow_low_2
  ) |>
  summarise(
    n_backgrounds = n_distinct(background_id),
    median_delta_pp =
      100 * median(delta_average_utilization, na.rm = TRUE),
    mean_delta_pp =
      100 * mean(delta_average_utilization, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    comparison_label = comparison_labels[comparison],
    noshow_low_2 = factor(noshow_low_2)
  )

noshow_dip_16 <- noshow_pattern |>
  filter(
    comparison == "both_flexible_vs_baseline",
    rho == 1.6,
    as.character(noshow_low_2) == "0.3"
  )

noshow_dip_20 <- noshow_pattern |>
  filter(
    comparison == "both_flexible_vs_baseline",
    rho == 2.0,
    as.character(noshow_low_2) == "0.3"
  )

ggplot(
  noshow_pattern,
  aes(
    x = rho,
    y = median_delta_pp,
    colour = noshow_low_2,
    group = noshow_low_2
  )
) +
  geom_hline(yintercept = 0, colour = "grey55") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    colour = "grey70"
  ) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  facet_wrap(~comparison_label, ncol = 2) +
  labs(
    x = "Demand-to-capacity ratio",
    y = "Median utilization difference (pp)",
    colour = paste(
      "General-group low-delay",
      "no-show probability"
    )
  ) +
  theme_report()
```

The lines connect grouped medians rather than following the same backgrounds as demand changes. For the 0.30 general-group no-show line in the both-flexible panel, the 1.6 point summarizes `r noshow_dip_16$n_backgrounds` backgrounds and the 2.0 point summarizes `r noshow_dip_20$n_backgrounds`. The medians move from `r paste0(formatC(noshow_dip_16$median_delta_pp, format = "f", digits = 2), " pp")` to `r paste0(formatC(noshow_dip_20$median_delta_pp, format = "f", digits = 2), " pp")`, while the means move from `r paste0(formatC(noshow_dip_16$mean_delta_pp, format = "f", digits = 2), " pp")` to `r paste0(formatC(noshow_dip_20$mean_delta_pp, format = "f", digits = 2), " pp")`. The visible dip is therefore mainly a median-and-composition effect, not evidence that raising demand from 1.6 to 2.0 reduces reservation value. Each point contains a different mixture of horizons, capacities, group shares, thresholds, cancellation rates, and balking behavior. The same caution applies to the reservation-only panel.

Under strict reservations:

- Horizon flexibility has little median effect until demand becomes very high.
- Reservation gains rise when the general group has a high low-delay no-show probability, especially once demand exceeds capacity.
- At a demand-to-capacity ratio of 3.0, the median reservation-only gain ranges from about **0.3 pp** when the general-group low-delay no-show probability is 0.05 to about **3.8 pp** when it is 0.30.
- The no-show thresholds and other behavioral parameters modify the result, but their patterns are weaker and less uniform than demand pressure.

## Optimized policy parameters

The figure below focuses on **both-flexible**, where the horizon, reserved quantity, and reservation window were jointly selected.

```{r}
#| label: fig-optimized-parameters
#| fig-cap: "Median jointly optimized parameters for the both-flexible policy. Lines connect different background groups and candidate values are discrete; they should not be read as smooth response curves."
#| fig-width: 10.5
#| fig-height: 7

joint_policy <- avg_outcomes |>
  filter(policy == "both_flexible") |>
  mutate(
    reserved_capacity_pct =
      100 * selected_Q / slots_per_day
  )

window_point_rho2_share03 <- joint_policy |>
  filter(
    rho == 2.0,
    class1_share == 0.3,
    selected_Q > 0,
    selected_window >= 1
  )

parameter_summary <- bind_rows(
  joint_policy |>
    group_by(rho, class1_share) |>
    summarise(
      value = median(
        selected_horizon_days,
        na.rm = TRUE
      ),
      metric = "Selected booking horizon (days)",
      .groups = "drop"
    ),
  joint_policy |>
    group_by(rho, class1_share) |>
    summarise(
      value = median(
        reserved_capacity_pct,
        na.rm = TRUE
      ),
      metric = "Reserved quantity (% of daily capacity)",
      .groups = "drop"
    ),
  joint_policy |>
    filter(
      selected_Q > 0,
      selected_window >= 0
    ) |>
    group_by(rho, class1_share) |>
    summarise(
      value = median(
        selected_window,
        na.rm = TRUE
      ),
      metric = "Reservation-window length (days)",
      .groups = "drop"
    )
) |>
  mutate(
    class1_share = factor(class1_share)
  )

ggplot(
  parameter_summary,
  aes(
    x = rho,
    y = value,
    colour = class1_share,
    group = class1_share
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  facet_wrap(
    ~metric,
    scales = "free_y",
    ncol = 1
  ) +
  labs(
    x = "Demand-to-capacity ratio",
    y = "Median selected parameter\n(days or % of daily capacity)",
    colour = "Priority-group share"
  ) +
  theme_report()
```

These lines also connect grouped medians over a discrete search grid; they are not smooth response curves. At a demand-to-capacity ratio of 2.0 and a priority-group share of 0.3, the six-day reservation-window point summarizes `r nrow(window_point_rho2_share03)` backgrounds with positive reservations. Their selected windows range from `r min(window_point_rho2_share03$selected_window)` to `r max(window_point_rho2_share03$selected_window)` days. Six days is therefore the middle of a wide distribution, not a uniquely preferred window for every 30% priority-group scenario. A plausible mechanism is that a moderate priority share sometimes needs a longer near-term window to expose enough Class 1 arrivals to reserved capacity, whereas a very small priority group makes reservations harder to fill and a large priority group can fill them within a shorter window. Because the remaining scenario parameters also vary, this figure does not isolate that mechanism.

The selected booking horizon can be six days because the optimization grid was **2, 6, 10, 14, 18, 22, and 26 days**. The study did not restrict optimized horizons to three weeks or longer. Under congestion, a short horizon can improve utilization by avoiding long-delay offers that trigger balking or higher no-show risk while still leaving enough demand to fill near-term capacity.

Strict reservation parameters are conservative because unused priority capacity cannot be returned to the general group:

- The jointly optimized horizon is **2 days** in `r pct(mean(joint_policy$selected_horizon_days == 2))` of backgrounds and **6 days** in `r pct(mean(joint_policy$selected_horizon_days == 6))`.
- Reserved quantity generally rises with the priority-group share.
- When the priority group is only 10% of arrivals, the median jointly selected reserved quantity is zero.
- For priority shares from 30% to 90%, the median reserved share rises approximately with the population share.
- Jointly optimized reservation windows are usually short—about two days—because long strict windows increase the risk of unused capacity.

## Effects on the two patient groups

The utilization objective counts completed appointments, regardless of which group receives them. A utilization gain can therefore coexist with a redistribution of access.

```{r}
#| label: fig-patient-group-effects
#| fig-cap: "Supported Class 1 and Class 2 served-rate patterns for each policy compared with baseline."
#| fig-width: 10
#| fig-height: 4.8

collapse_tradeoff <- function(x) {
  case_when(
    x == "both_groups_gain" ~
      "Both groups gain",
    x == "priority_gains_general_loses" ~
      "Priority gains; general loses",
    x == "priority_loses_general_gains" ~
      "General gains; priority loses",
    x %in% c(
      "priority_gains_general_tied",
      "priority_tied_general_gains"
    ) ~
      "One group gains; other tied",
    x %in% c(
      "priority_loses_general_tied",
      "priority_tied_general_loses",
      "both_groups_lose"
    ) ~
      "At least one group loses",
    TRUE ~
      "No clear group change"
  )
}

group_effects <- avg_pairwise |>
  filter(
    comparison %in% c(
      "horizon_only_vs_baseline",
      "reservation_only_vs_baseline",
      "both_flexible_vs_baseline"
    )
  ) |>
  mutate(
    comparison_label = comparison_labels[comparison],
    group_pattern = collapse_tradeoff(
      patient_group_supported_tradeoff
    )
  ) |>
  count(
    comparison_label,
    group_pattern,
    name = "n"
  ) |>
  group_by(comparison_label) |>
  mutate(share = n / sum(n)) |>
  ungroup()

group_pattern_levels <- c(
  "Both groups gain",
  "Priority gains; general loses",
  "General gains; priority loses",
  "One group gains; other tied",
  "At least one group loses",
  "No clear group change"
)

group_effects <- group_effects |>
  mutate(
    group_pattern = factor(
      group_pattern,
      levels = group_pattern_levels
    ),
    comparison_label = factor(
      comparison_label,
      levels = c(
        "Both flexible − baseline",
        "Reservation only − baseline",
        "Horizon only − baseline"
      )
    )
  )

ggplot(
  group_effects,
  aes(
    x = share,
    y = comparison_label,
    fill = group_pattern
  )
) +
  geom_col(width = 0.7) +
  scale_x_continuous(labels = percent_format()) +
  scale_fill_manual(
    values = c(
      "#2A9D8F",
      "#E76F51",
      "#457B9D",
      "#F4A261",
      "#9D4EDD",
      "#D9D9D9"
    ),
    drop = FALSE
  ) +
  labs(
    x = "Share of backgrounds",
    y = NULL,
    fill = "Supported group pattern"
  ) +
  theme_report()
```

- Horizon changes can help or hurt either group because they change the set of delays offered to everyone.
- Reservations more directly redistribute access toward Class 1.
- “Priority gains; general loses” means the paired confidence intervals support both directions. It is not inferred from point estimates alone.

The main strict-reservation trade-off appears under congested conditions:

- Reservation-only produces a supported priority-gain/general-loss pattern in `r pct(tradeoff_rate("reservation_only_vs_baseline"))` of backgrounds.
- Both-flexible produces the same pattern in `r pct(tradeoff_rate("both_flexible_vs_baseline"))`.
- At demand ratios below 1.2, clear group changes are rare.
- As demand rises, reservations increasingly determine **who** receives limited capacity rather than creating additional capacity.

## Comparison with the twice-weighted Class 1 objective

The secondary objective was:

> **Priority-weighted served rate** = [2 × **Class 1 served rate** + **Class 2 served rate**] ÷ 3

Class 1 therefore receives **twice the policy weight** of Class 2. The code stores this metric as `weighted_utilization`.

### What is similar?

- The baseline is identical because it has no parameters to optimize.
- Demand remains the main condition under which policy changes matter.
- Short horizons remain common under both objectives.
- Horizon-only is more stable across objectives than reservation policies.

### What changes?

```{r}
#| label: tbl-objective-switch

objective_switch_summary <- objective_switch |>
  filter(policy != "baseline") |>
  group_by(policy) |>
  summarise(
    `Same exact policy cell` =
      mean(same_policy_cell, na.rm = TRUE),
    `Median average-utilization change` =
      median(delta_average_utilization, na.rm = TRUE),
    `Median weighted-rate change` =
      median(delta_weighted_utilization, na.rm = TRUE),
    `Average-utilization loss > 0.5 pp` =
      mean(delta_average_utilization < -0.005, na.rm = TRUE),
    `Weighted-rate gain > 0.5 pp` =
      mean(delta_weighted_utilization > 0.005, na.rm = TRUE),
    `Priority gains; general loses` =
      mean(
        patient_group_supported_tradeoff ==
          "priority_gains_general_loses",
        na.rm = TRUE
      ),
    .groups = "drop"
  ) |>
  mutate(
    Policy = policy_labels[policy],
    `Same exact policy cell` =
      pct(`Same exact policy cell`),
    `Median average-utilization change` =
      pp(`Median average-utilization change`),
    `Median weighted-rate change` =
      pp(`Median weighted-rate change`),
    `Average-utilization loss > 0.5 pp` =
      pct(`Average-utilization loss > 0.5 pp`),
    `Weighted-rate gain > 0.5 pp` =
      pct(`Weighted-rate gain > 0.5 pp`),
    `Priority gains; general loses` =
      pct(`Priority gains; general loses`)
  ) |>
  select(
    Policy,
    `Same exact policy cell`,
    `Median average-utilization change`,
    `Median weighted-rate change`,
    `Average-utilization loss > 0.5 pp`,
    `Weighted-rate gain > 0.5 pp`,
    `Priority gains; general loses`
  )

kable(
  objective_switch_summary,
  align = c("l", rep("r", 6))
)
```

```{r}
#| label: fig-objective-tradeoff
#| fig-cap: "Changing from the average-utilization optimum to the twice-weighted Class 1 optimum. Upper-left observations gain priority-weighted access while losing average utilization."
#| fig-width: 10
#| fig-height: 4.8

switch_plot <- objective_switch |>
  filter(policy != "baseline") |>
  mutate(
    policy_label = policy_labels[policy],
    average_change_pp =
      100 * delta_average_utilization,
    weighted_change_pp =
      100 * delta_weighted_utilization
  )

ggplot(
  switch_plot,
  aes(
    x = average_change_pp,
    y = weighted_change_pp
  )
) +
  geom_hline(yintercept = 0, colour = "grey45") +
  geom_vline(xintercept = 0, colour = "grey45") +
  geom_hline(
    yintercept = 0.5,
    linetype = "dotted",
    colour = "grey65"
  ) +
  geom_vline(
    xintercept = -0.5,
    linetype = "dotted",
    colour = "grey65"
  ) +
  geom_point(
    alpha = 0.24,
    size = 1.2
  ) +
  facet_wrap(
    ~policy_label,
    scales = "free",
    ncol = 3
  ) +
  labs(
    x = "Change in average utilization (pp)",
    y = "Change in priority-weighted served rate (pp)"
  ) +
  theme_report()
```

The weighted objective mainly changes reservation design:

- For reservation-only, switching to the weighted optimum changes average utilization by a median of `r pp(median_switch_average("reservation_only"))` and raises the weighted served rate by `r pp(median_switch_weighted("reservation_only"))`.
- For both-flexible, the corresponding medians are `r pp(median_switch_average("both_flexible"))` and `r pp(median_switch_weighted("both_flexible"))`.
- The trade-off is strongest when the priority group is small or moderate and demand is high. More capacity can be redirected to Class 1, but strict unused reservations cannot be recovered.
- Horizon-only is much more similar across objectives because it changes access timing for both groups rather than reserving capacity for one group.

## Brief comparison with the companion report

The [same-day-release report](h1_release_policy_comparison.html) shows the same broad demand pattern, but release changes reservation behavior:

- Reservation-only has a supported utilization increase in `r pct(supported_rate("reservation_only_vs_baseline"))` of strict backgrounds versus `r pct(mean(avg_companion_pairwise |> filter(comparison == "reservation_only_vs_baseline") |> pull(average_utilization_change_status) == "supported_increase"))` under release.
- Release permits much larger nominal reservation quantities because unused slots can return to the general pool.
- The weighted-access objective imposes a smaller average-utilization penalty under release than under strict reservation.

## Conclusions

- **Low and moderate demand:** practical ties dominate; policy flexibility usually adds little average utilization.
- **Heavy demand:** short horizons drive the largest gains. Reservations can add further value when the general group has high no-show risk.
- **Single-policy comparison:** reservation-only and horizon-only each win in some backgrounds; most are tied.
- **Access distribution:** strict reservations frequently raise priority-group access while reducing general-group access.
- **Objective choice matters:** giving Class 1 twice the policy weight often changes `Q` and the reservation window and can reduce total utilization.
